Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-14 PAPER.tex
-Made Danny's edits for first half of paper

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to develop techniques that help maximize sensitivity and validate results. Additionally, it is imperative to understand the trade-offs between different methods and their effects on common power spectrum themes. In this paper, we focus on three major themes --- signal loss, power spectrum error bar estimation, and bias in measurements. We describe techniques that affect these themes using both a toy model and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have led to a revised, higher $21$ cm power spectrum upper limit from PAPER-64. 
\end{abstract}

@@

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and instrumental systematics that are, when combined, 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure the accuracy and reliability of a potential EoR detection. In this paper, we discuss three themes that are essential to investigate for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, traps, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed case study using data from the 64-element configuration of PAPER.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power spectrum result. Stemmed from a re-analysis of PAPER-64 data, we believe these themes serve as an important check-list for a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): As explained in the next section, certain analysis techniques can lead to the loss of the EoR signal. If not corrected for, it could lead to a false non-detection. Computing signal loss correctly has subtle challenges but is necessary to ensure the accuracy of any result.
\item \textbf{Error Bar Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result determine the difference between a detection and a null result, which have two very different implications. Errors can be estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of bias in a visibility measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a successful EoR detection would also imitate a bias. Proving a bias is EoR may be the most difficult challenge for $21$ cm analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper we will highlight some sources of bias, discuss ways to mitigate their effects, and describe tests that a true EoR detection must pass.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also being standard steps in general $21$ cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is considered an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}), and we explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}. Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. 
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which is a generalized version of inverse variance weighting that also takes into account data correlations. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, a challenge of inverse covariance weighting is in accurately describing a covariance matrix that best describes the data. 
\item \textbf{Bootstrapping:} Because theoretical models for covariance matrices are questionable and theoretical error estimation methods rely on assumptions that can be violated by data, applying theoretical errors to a $21$ cm measurement is often avoided. Hence, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we obtain a sense of its inherent variance, though there are subtleties to consider such as the independence of values in a dataset.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}). An EoR detection must pass jack-knife and null tests.
\end{itemize}

@@

In the next sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds perfectly without destroying an EoR signal.

@@

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid that varies in time and frequency, a simplistic but realistic representation of a single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal \cc{need to motivate why, since real EoR does have correlations}. Our combined data vector is $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

@@

One choice for the weighting matrix $\textbf{R}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds and yields the smallest possible error bars on a measurement (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). The covariance matrix $\textbf{C}$, which in our case describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting them in the data. For PAPER, we are mostly limited from being able to estimate a true $\textbf{C}$ due to our instrumental models.

@@

It turns out that because we estimated $\hat{\textbf{C}}$ from our data, its eigenspectrum differs from the eigenspectrum of our true $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from highest to lowest and can be thought of as a spectrum of weights that are given to each frequency mode in the data. In other words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, in which a sloped eigenspectrum means that modes are given different weights. When weighting data by an inverse covariance matrix, the modes with the highest eigenvalues are down-weighted the most.

@@

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are more trustworthy than the weakest modes. Bright foregrounds usually dominate and would therefore exist in both $\textbf{C}$ and $\hat{\textbf{C}}$. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid and assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}).

The danger of an empirically estimated covariance matrix comes from not being able to describe weak eigenmodes accurately. The weak eigenmodes of $\hat{\textbf{C}}$ may characterize random noise fluctuations which will be down-weighted. This is what we call the `overfitting' of noise, which leads to signal loss.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Namely, we can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Hence, we are not using information from all the weaker eigenmodes which carry the risk of describing fluctuations that may not exist in the true covariance. Altering $\hat{\textbf{C}}$ as such is one example of a regularization method, in which we are changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more identical to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds perfectly (and disentangle them from other modes), we can down-weight them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

A fringe-rate filter is an analysis technique designed to maximize sensitivity by integrating in time (\citealt{parsons_et_al2016}). Rather than a traditional box-car average, a fringe-rate filter can be designed to up-weight parts of the sky that an instrument is most sensitive to, while down-weighting the opposite. We apply a fringe-rate filter to PAPER-64 data in order to optimally combine our time-ordered measurements. We weight data based on our primary beam sensitivity, yielding a net increase in sensitivity in our measurements.

@@

The resulting eigenspectra (Figure \ref{fig:toy_sigloss5}, right), as compared to those in Figure \ref{fig:toy_sigloss2}, fall more steeply, especially for the last few eigenmodes. This is because we have fewer independent modes --- fewer shapes in the data --- so we do a worse job characterizing the weak eigenmodes. Therefore, there is a more dramatic difference in weighting between the modes and we end up down-weighting random noise-like fluctuations more severely than for the un-fringe-rate filtered case.

\begin{figure*}
	\centering
	\includegraphics[trim={1.5cm 0.2cm 3.5cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss6.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of $\hat{\textbf{C}}_{FG}$ (blue), $\hat{\textbf{C}}_{EoR}$ (red), and $\hat{\textbf{C}}_{FG+EoR}$ (green). The eigenspectrum of foregrounds only peaks at the zeroth eigenmode, a consequence of the presence of only one sinusoid. This spectrum is steeper than that in Figure \ref{fig:toy_sigloss2}, resulting in more dramatic differences in weighting between eigenmodes. Consequently, if all eigenmodes are used in weighting it is possible to down-weight random noisy modes more severely than for the un-fringe-rate filtered case.}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data. There is a larger amount of signal loss than for un-fringe-rate filtered data, a consequence of a steepened, weakly characterized eigenspectrum from using fewer independent modes in the data. This example shows that it is possible to down-weight modes more severely for the un-fringe-rate filtered since there is a more dramatic difference in weighting between modes. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes.}
	\label{fig:toy_sigloss7}
\end{figure}

@@

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal loss and noisier estimates when using data-estimated covariance matrices. Longer integrations increase sensitivity but reduce the number of independent samples, resulting in weakly characterized, steep eigenspectra that can overfit noise greatly.

@@

As a first test, we know that our simulated EoR should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. If we model $\textbf{C}_{EoR}$ as such, instead of computing it based on the dataset itself, and add it to $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$ to obtain a final $\hat{\textbf{C}}$ to use in weighting, we see that there is negligible signal loss (Figure \ref{fig:toy_sigloss8}, upper left). This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting small fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves differ at one $k$ value in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices.

\begin{figure*}
	\centering
	\includegraphics[trim={0.4cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={1cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0.4cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={1cm 0.8cm 1.3cm 1cm},clip,height=0.25\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each avoid signal loss, including modeling the covariance matrix of EoR (upper left), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to it (upper right), using only the first three eigenmodes of $\hat{\textbf{C}}$ (lower left), and multiplying an identity matrix with $\hat{\textbf{C}}$ (lower right).}
	\label{fig:toy_sigloss8}
\end{figure*}

@@

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}}_{ij} + \textbf{I}_{ij}$, where $i$$=$$j$$=$$20$ (number of frequencies). By adding the identity matrix, element-wise, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum. If all modes are given similar weights, we avoid down-weighting artificial modes more than others.

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} flattens out the eigenspectrum of $\hat{\textbf{C}}$ a different way - by zeroing all but the first three eigenmodes. This means that the first three eigenmodes will be given weights, but the rest will be treated with similar weights. Again, flattening the eigenspectrum results in negligible signal loss. However, we do not perfectly recover the shape of EoR because we lost information when projecting out certain modes.

The last regularization scheme we are highlighting here is setting $\hat{\textbf{C}} = \hat{\textbf{C}}_{ij}\textbf{I}_{ij}$, or inverse variance weighting. In the bottom right panel of Figure \ref{fig:toy_sigloss8}, we see that this method does a poor job down-weighting foregrounds. For this toy model, our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by the identity, element-wise, results in a diagonal matrix, meaning we are only left with correlation information between the same two frequencies. Because we disregard information from all other frequency combinations in this case, we do a poor job suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate confidence intervals for our measurements. Two popular ways of estimating errors on a power spectrum measurement are calculating the variance of a dataset, and computing a theoretical error estimate based on an instrument's system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the two don't always agree due to a number of factors, including time, frequency, and antenna-dependent noise and non-uniform weightings. Therefore, it is important to place error bars on our measurements that have been derived from its inherent variance.

A common technique used to estimate the error in a measurement is bootstrapping. Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, power-spectral measurements of $21$ cm data can be made along many axes, including time and baselines. Through the process of resampling and averaging along these axes, we can estimate error bars for our results which represent the underlying distribution of power spectra values that are allowed by our measurements.

@@

This examples highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the number of averages performed. We can also have a non-integer effective number of samples due to the weighting. Hence, we do not recommend bootstrapping unless the number of independent samples along the axis that is being re-sampled is well-determined.

>>>>

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the birth of the first stars and galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to develop techniques that help maximize sensitivity and validate results. Additionally, it is imperative to understand the trade-offs between different methods and their effects on common power spectrum themes. In this paper, we focus on three major themes --- signal loss, power spectrum error bar estimation, and bias in measurements. We describe techniques that affect these themes using both a toy model and data taken by the 64-element configuration of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have led to a revised, higher $21$ cm power spectrum upper limit from PAPER-64. This revised result mostly stems from an updated signal loss calculation for loss associated with empirically estimated covariances, and we consider this limit to supersede previously published PAPER-64 and PAPER-32 results.
\end{abstract}

@@

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and instrumental systematics that are, when combined, 4-5 orders of magnitude brighter. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure the accuracy and reliability of a potential EoR detection. In this paper, we discuss three themes that are essential to investigate for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, traps, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 $21$ cm power spectrum from the lessons learned.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power spectrum result. Stemmed from a re-analysis of PAPER-64 data, we believe these themes serve as an important check-list for a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): As explained in the next section, certain analysis techniques can lead to the loss of the EoR signal. If not corrected for, it could lead to a false non-detection. Computing signal loss correctly has subtle challenges but is necessary to ensure the accuracy of any result. PAPER's analysis is not the first to underestimate loss; results from the GMRT (\citealt{paciga_et_al2013}) were also revised from new signal loss calculations associated with their singular value decomposition foreground filter.
\item \textbf{Error Bar Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result determine the difference between a detection and a null result, which have two very different implications. Errors can be estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of positive power offset in a visibility measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a successful EoR detection would also imitate a bias. Proving a bias is EoR may be the most difficult challenge for $21$ cm analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper we will highlight some sources of bias, discuss ways to mitigate their effects, and describe tests that a true EoR detection must pass.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also being standard steps in general $21$ cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data (\citealt{parsons_et_al2016}), and we explain the trade-offs of filtering in more detail in Section \ref{sec:SiglossOverview}. Broadly, a fringe-rate filter increases the sensitivity of a dataset and reduces the number of independent samples by an amount dependent on the width of the averaging window. However, it can also affect the presence of foregrounds and systematics. 
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain spectral features and minimize others. One particular flavor of weighting is inverse covariance weighting, which is a generalized version of inverse variance weighting that also takes into account data correlations. This weighting has the effect of down-weighting correlated information (i.e. foregrounds) and up-weighting noise-like information (i.e. EoR). However, a challenge of inverse covariance weighting is in accurately defining a covariance matrix that best describes the data. 
\item \textbf{Bootstrapping:} Because theoretical models for covariance matrices are questionable and theoretical error estimation methods rely on assumptions that can be violated by data, applying theoretical errors to a $21$ cm measurement is often avoided. Hence, bootstrapping is a useful method for estimating errors of a dataset from itself. By randomly drawing many samples of the data, we obtain a sense of its inherent variance, though there are subtleties to consider such as the independence of values in a dataset.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of systematics (\citealt{keating_et_al2016}). An EoR detection must pass jack-knife and null tests.
\end{itemize}

@@

In the next sections, we investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examine their impact on the resulting power spectra estimate $\hat{\textbf{p}}$. Our ultimate goal in experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it.

@@

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which might be used for a power spectrum analysis. For PAPER-64 we use $\sim$$8$ hours of data for our analysis, but here we scale it down for this toy model with no loss of generality. We create a mock bright foreground signal, $\textbf{x}_{FG}$, as a complex sinusoid with a fixed frequency and a smoothly varying offset in time, a simplistic but realistic representation of a single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal \cc{need to motivate why, since real EoR does have correlations}. Our combined data vector is $\textbf{x} = \textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure \ref{fig:toy_sigloss1}.

@@

One choice for the weighting matrix $\textbf{R}$ is an inverse covariance matrix. This type of weighting is attractive for power spectrum analyses because it is an aggressive way to down-weight foregrounds and yields the smallest possible error bars on a measurement (\citealt{tegmark_et_al1997a}; \citealt{bond_et_al1998}). The covariance matrix $\textbf{C}$, which in our case describes covariances between frequency channels, can be estimated in a variety of ways. Given perfect foreground, instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data. However, if our foreground model is flawed for example, our estimate of $\textbf{C}$ would not be successful at down-weighting them in the data.

@@

It turns out that because we estimated $\hat{\textbf{C}}$ from our data, its eigenspectrum differs from the eigenspectrum of our true $\textbf{C}$, and this difference has consequences on our result. An eigenspectrum ranks the eigenvalues of a matrix from highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. This is usually not the case for a covariance matrix, in which a sloped eigenspectrum means that modes are given different weights. When weighting data by an inverse covariance matrix, the modes with the highest eigenvalues are down-weighted the most.

@@

In general, the strongest modes of $\hat{\textbf{C}}$ (highest eigenvalues) are more trustworthy than the weakest modes, since that's where we expect bright foregrounds --- the most prominent shapes --- to be. If bright foregrounds dominate, these highest eigenvalue modes therefore exist in both $\textbf{C}$ and $\hat{\textbf{C}}$. This is demonstrated in the toy model by the successful suppression of the foreground mode, where our estimated covariance matrix identifies the sinusoid and assigns it the highest eigenvalue (the peak in Figure \ref{fig:toy_sigloss2}).

The danger of an empirically estimated covariance matrix comes from not being able to describe weak eigenmodes of $\textbf{C}$ accurately. Instead, the weak eigenmodes of $\hat{\textbf{C}}$ may characterize noise-like fluctuations which will be down-weighted. These noise fluctuations could be the signal itself, since $\hat{\textbf{C}}$ is constructed using data that contains the signal. This leads to the `overfitting' of noise (or `overfitting' of signal), which is signal loss.

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero signal loss. Namely, we can project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds perfectly and nothing else. Hence, we are not using information from all the weaker eigenmodes which carry the risk of describing fluctuations that may not exist in the true covariance. Altering $\hat{\textbf{C}}$ as such is one example of a regularization method, in which we are changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more similar to that of $\textbf{I}$. The resulting power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, demonstrating that if we can model our foregrounds perfectly (and disentangle them from other modes), we can down-weight them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section \ref{sec:otherweight}.

@@

A fringe-rate filter is an analysis technique designed to maximize sensitivity by integrating in time (\citealt{parsons_et_al2016}). Rather than a traditional box-car average, a time domain filter can be designed to up-weight temporal modes relevant to things that move as expected for the sidereal motion on the sky, while down-weighting modes which are noise-like.

@@

The resulting eigenspectrum as compared to the green curve (FG + EoR) in Figure \ref{fig:toy_sigloss2} is shown in Figure \ref{fig:toy_sigloss5} (right plot). The spectrum, in the case of fringe-rate filtering (dashed line), falls more steeply, especially for the last few eigenmodes. This is because we have fewer independent modes --- \cc{need help filling out the rest here... and also citing galaxy lit}.

\begin{figure*}
	\centering
	\includegraphics[trim={1.5cm 0.2cm 3.5cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss5.png}
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,height=0.3\textwidth]{plots/toy_sigloss15.png}
	\caption{Left: Our `fringe-rate filtered' (time-averaged) toy model dataset. We average every four samples together, yielding $25$ independent samples in time. Real parts are shown here. Right: Eigenspectrum of $\hat{\textbf{C}}_{FG+EoR}$, in the case of no fringe-rate filtering (solid green) and with fringe-rate filtering (dashed green). The dashed spectrum is steeper, resulting in more dramatic differences in weighting between eigenmodes. Consequently, if all eigenmodes are used in weighting it is possible to down-weight random noisy modes more severely than for the non-averaged case.}
	\label{fig:toy_sigloss5}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.2cm 0.3cm 0.3cm},clip,width=\columnwidth]{plots/toy_sigloss7.png}
	\caption{Resulting power spectrum estimate for the fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We use inverse covariance weighting where $\textbf{C}$ is derived from the data. There is a larger amount of signal loss than for the non-averaged data, a consequence of a steepened, weakly characterized eigenspectrum from using fewer independent modes in the data. The power spectrum estimate is also noisier, since most of the information used in constructing the estimate came from only a few eigenmodes.}
	\label{fig:toy_sigloss7}
\end{figure}

@@

Using our toy model, we have seen that a sensitivity-driven analysis technique like fringe-rate filtering has trade-offs of signal loss and noisier estimates when using data-estimated covariance matrices. Longer integrations increase sensitivity but reduce the number of independent samples, resulting in weakly characterized, steep eigenspectra that can overfit noise greatly. \cc{help me fix explanation}

@@

As a first test, we know that our simulated EoR should have a covariance matrix that mimics the identity matrix, with its variance encoded along the diagonal. If we model $\textbf{C}_{EoR}$ as such, instead of computing it based on the dataset itself, and add it to $\hat{\textbf{C}}_{FG}$ (defined as $\hat{\textbf{C}}_{FG} = \langle\textbf{x}_{FG}\textbf{x}_{FG}^{\dagger}\rangle$) to obtain a final $\hat{\textbf{C}}$ to use in weighting, we see that there is negligible signal loss (Figure \ref{fig:toy_sigloss8}, upper left). This is because by modeling $\textbf{C}_{EoR}$, we avoid over-fitting small fluctuations in the data that our model doesn't know about (but an empirically derived $\hat{\textbf{C}}$ would). This is evident when comparing the green and red curves in Figure \ref{fig:toy_sigloss14}. We don't see perfect EoR recovery though, as the green and red curves differ at one $k$ value in Figure \ref{fig:toy_sigloss8}. This deviation is a consequence of the difference between the true and modeled covariance matrices.

\begin{figure*}
	\centering
	\includegraphics[trim={0.4cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss10.png}
	\includegraphics[trim={1cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss8.png}
	\includegraphics[trim={0.4cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss9.png}
	\includegraphics[trim={1cm 0cm 1.3cm 0cm},clip,height=0.3\textwidth]{plots/toy_sigloss11.png}
	\caption{Resulting power spectra estimates for our fringe-rate filtered toy model simulation --- foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). We show four alternate weighting options that each avoid signal loss, including modeling the covariance matrix of EoR (upper left), regularizing $\hat{\textbf{C}}$ by adding an identity matrix to it (upper right), using only the first three eigenmodes of $\hat{\textbf{C}}$ (lower left), and keeping only the diagonal elements of $\hat{\textbf{C}}$ (lower right).}
	\label{fig:toy_sigloss8}
\end{figure*}

@@

The second panel (top right) in Figure \ref{fig:toy_sigloss8} uses a regularization method of setting $\hat{\textbf{C}} = \hat{\textbf{C}}_{ij} + \gamma\textbf{I}_{ij}$, where $i$$=$$j$$=$$20$ (number of frequencies) and $\gamma = 5$ (an arbitrary strength of $\textbf{I} for the purpose of this toy model$). By adding the identity matrix, element-wise, we are weighting the diagonal elements of the matrix more heavily than those off-diagonal, thereby flattening out its eigenspectrum. If all modes are given similar weights, we avoid down-weighting artificial modes more than others.

The third panel (bottom left) in Figure \ref{fig:toy_sigloss8} flattens out the eigenspectrum of $\hat{\textbf{C}}$ a different way - by only using the first three eigenmodes. This means that all but the three highest weights will be replaced with $1$'s (equal weights). Again, flattening the eigenspectrum results in negligible signal loss. However, we do not perfectly recover the shape of EoR because we lost information when projecting out certain modes.

The last regularization scheme we are highlighting here is setting $\hat{\textbf{C}} = \hat{\textbf{C}}_{ij}\textbf{I}_{ij}$, or inverse variance weighting (keeping only the diagonal elements of $\hat{\textbf{C}}$). In the bottom right panel of Figure \ref{fig:toy_sigloss8}, we see that this method does a poor job down-weighting foregrounds. For this toy model, our foregrounds are spread out in frequency and therefore have non-negligible frequency-frequency correlations. Multiplying by the identity, element-wise, results in a diagonal matrix, meaning we are only left with correlation information between the same two frequencies. Because we disregard information from all other frequency combinations in this case, we do a poor job suppressing the foreground. But because we flattened the eigenspectrum, we also avoid signal loss.

@@

Our second major $21$ cm power spectrum theme is error estimation, as we desire robust methods for determining accurate confidence intervals for our measurements. Two popular ways of estimating errors on a power spectrum measurement are calculating the variance of a dataset, and computing a theoretical error estimate based on an instrument's system temperature and observational parameters. In a perfect world, both methods would match up. However, in practice the two don't always agree due to a number of factors, including time, frequency, and antenna-dependent noise and non-uniform weightings. Since modeling errors for a complex instrument is only approximate, we choose to compute error bars on our measurements that have been derived from its inherent variance.

A common technique used to estimate the error in a measurement is bootstrapping. Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, power-spectral measurements of $21$ cm data can be made along many axes, including time and baselines. Averaging each sample set gives us a realization of the power spectrum answer. Realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common between them. Hence, through the process of resampling and averaging along these axes, we can estimate error bars for our results which represent the underlying distribution of power spectra values that are allowed by our measurements.

@@

This example highlights the importance of understanding how analysis techniques (e.g. fringe-rate filtering) can affect a common statistical procedure like bootstrapping. Bootstrapping as a means of estimating power spectrum errors from real fringe-rate filtered data requires knowledge of the number of independent samples, which is not always a trivial task. For example, computing the effective number of independent samples of fringe-rate filtered data is not as simple as counting the number of averages performed. Down-sampling a fringe-rate filtered signal is straightforward using a boxcar average, but non-trivial with a more complicated convolution function that has long tails. Hence, we do not recommend bootstrapping unless the number of independent samples along the axis that is being re-sampled is well-determined.

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error1.png}
	\caption{Error estimation produced by bootstrapping as a function of the number of samples drawn when sampling with replacement. The star represents the error associated with drawing $1000$ samples from a length $1000$ array of a Gaussian random signal. The black points correspond to time-averaged data (correlated data) which has $100$ independent samples. They illustrate how errors can be underestimated if drawing more samples than there are independent samples in the data. The estimated errors match up with the theoretical prediction only at $N=100$.}
	\label{fig:toy_error1}
\end{figure}
